ISYE 6501

### 2022-10-19 hw7/submission-hw7.Rmd
backlog

2. Another point to note is that the Multiple R-squared of the final model is 0.789 which is lower than the initial model which has a value of 0.8031 but this is the pitfall of R squared value i.e. It will always increases with the increase in the # of variables without any penalty/adjustment for model complexity/overfitting unlike AIC or Adjusted R-squared.

@@

Based on the above formula , we can infer that the Adjusted R squared value decreases with the increase in # of independent variables(k) until and unless if there is a significant increase in the R squared value on adding a new variable. Thus both AIC and Adjusted R squared provide a technique for selecting simpler model which prevents overfitting.

@@

@@

Two special values along the 𝜆 sequence are indicated by the vertical dotted lines. `lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

@@

@@

@@

### RMSE for

@@

* Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
* It involves two hyperparameters - Lambda and Alpha.
* Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
* Lambda is amount of penalty in the regression.

@@

* Splitting the data to get a value of MSE for deciding a value of alpha.

@@

* We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
* We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical. We are accepting some performance loss and incorporating more bias in the model.

@@

* After we have selected our lambda, we predict on the test data for getting the value of alpha.
* Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.

@@

--------  DO NOT INCLUDE AFTER THIS LINE ---------

```{r}
library(MASS)
library(caret)
```

```{r}
custom <- trainControl(method = "repeatedcv", number = 10, repeats = 5, verboseIter = TRUE)
```

```{r}
set.seed(1234)
en <- train(x = x.train, y = y.train,
            method='glmnet',
            tuneGrid =expand.grid(alpha=seq(0,1,length=10),
                                  lambda = seq(0.0001,0.2,length=5)),
            trControl=custom)
```

```{r}
en
```

```{r}
mean(en$resample$RMSE)
```

```{r}
plot(en, main = "Elastic Net Regression")
```

```{r}
plot(varImp(en,scale=TRUE))
```

>>>>

2.  Another point to note is that the Multiple R-squared of the final model is 0.789 which is lower than the initial model which has a value of 0.8031 but this is the pitfall of R squared value i.e. It will always increases with the increase in the \# of variables without any penalty/adjustment for model complexity/overfitting unlike AIC or Adjusted R-squared.

@@

Based on the above formula , we can infer that the Adjusted R squared value decreases with the increase in \# of independent variables(k) until and unless if there is a significant increase in the R squared value on adding a new variable. Thus both AIC and Adjusted R squared provide a technique for selecting simpler model which prevents overfitting.

@@

Lasso Regression is fit with glmnet function. Here we have two parameters:

-    alpha = 1 gives 1 weight to Lasso and 0 to Ridge

-   lambda is another parameter is the weight for regularization in cost function

Note the parameter standardizes the data, which is a requirement for Lasso

@@

Two special values along the λ sequence are indicated by the vertical dotted lines. `lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

@@

Let us fit full models for these parameter values. While ideally we should pick model from CV stage, just to understand the functioning better we are selecting two models here.

The nuance here is that while lambda.min minimizes CV error, We employ lasso for feature selection and hence one might argue lambda.1se is the parameter we should choose, since it minimizes the number of features selected.

@@

We also note here that the variables being dropped are similar to what we hypothesized in previous iterations of this exercise, say in RF and Linear Regression. With highly correlated features being dropped.

@@

### RMSE for above two models

@@

-   Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
-   It involves two hyperparameters - Lambda and Alpha.
-   Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
-   Lambda is amount of penalty in the regression.

@@

-   Splitting the data to get a value of MSE for deciding a value of alpha.

@@

-   We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
-   We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical. We are accepting some performance loss and incorporating more bias in the model.

@@

-   After we have selected our lambda, we predict on the test data for getting the value of alpha.
-   Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.

@@
